### 2019-05-05 InfrastructureMaintenance/Programs/IMP-bayesian/IMP-BNextendedexample.tex
Add files via upload

Bayesian network parameters
  Parameters of node Aval (multinomial distribution)
Conditional probability table:
, , Snow = high
     Earthquake
Aval         no       yes
  no  0.7500000 0.5000000
  yes 0.2500000 0.5000000
, , Snow = low
     Earthquake
Aval         no       yes
  no  0.6000000 0.6666667
  yes 0.4000000 0.3333333
  Parameters of node Bearing (multinomial distribution)
Conditional probability table:
       DTV
Bearing      high       low    medium
   bad  0.7000000 0.1666667 0.8571429
   good 0.3000000 0.8333333 0.1428571
  Parameters of node Bridge (multinomial distribution)
Conditional probability table:
, , Bearing = bad, Cracks = high, Flood = no, Joint = no, Rockfall = no
        Aval
Bridge   no yes
  high      1.0
  low       0.0
  medium    0.0
, , Bearing = good, Cracks = high, Flood = no, Joint = no, Rockfall = no
        Aval
Bridge    no yes
  high   0.5    
  low    0.0    
  medium 0.5    
, , Bearing = bad, Cracks = moderate, Flood = no, Joint = no, Rockfall = no
        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = good, Cracks = moderate, Flood = no, Joint = no, Rockfall = no

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = bad, Cracks = small, Flood = no, Joint = no, Rockfall = no

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = good, Cracks = small, Flood = no, Joint = no, Rockfall = no

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = bad, Cracks = high, Flood = yes, Joint = no, Rockfall = no

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = good, Cracks = high, Flood = yes, Joint = no, Rockfall = no

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = bad, Cracks = moderate, Flood = yes, Joint = no, Rockfall = no

        Aval
Bridge   no yes
  high      1.0
  low       0.0
  medium    0.0

, , Bearing = good, Cracks = moderate, Flood = yes, Joint = no, Rockfall = no

        Aval
Bridge    no yes
  high   0.5    
  low    0.0    
  medium 0.5

, , Bearing = bad, Cracks = small, Flood = yes, Joint = no, Rockfall = no

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = good, Cracks = small, Flood = yes, Joint = no, Rockfall = no

        Aval
Bridge   no yes
  high      1.0
  low       0.0
  medium    0.0

, , Bearing = bad, Cracks = high, Flood = no, Joint = yes, Rockfall = no

        Aval
Bridge    no yes
  high   0.0    
  low    1.0    
  medium 0.0

, , Bearing = good, Cracks = high, Flood = no, Joint = yes, Rockfall = no

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = bad, Cracks = moderate, Flood = no, Joint = yes, Rockfall = no

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = good, Cracks = moderate, Flood = no, Joint = yes, Rockfall = no

        Aval
Bridge   no yes
  high      1.0
  low       0.0
  medium    0.0

, , Bearing = bad, Cracks = small, Flood = no, Joint = yes, Rockfall = no

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = good, Cracks = small, Flood = no, Joint = yes, Rockfall = no

        Aval
Bridge    no yes
  high   1.0    
  low    0.0    
  medium 0.0

, , Bearing = bad, Cracks = high, Flood = yes, Joint = yes, Rockfall = no

        Aval
Bridge    no yes
  high   0.0    
  low    1.0    
  medium 0.0

, , Bearing = good, Cracks = high, Flood = yes, Joint = yes, Rockfall = no

        Aval
Bridge   no yes
  high      1.0
  low       0.0
  medium    0.0

, , Bearing = bad, Cracks = moderate, Flood = yes, Joint = yes, Rockfall = no

        Aval
Bridge    no yes
  high   1.0    
  low    0.0    
  medium 0.0

, , Bearing = good, Cracks = moderate, Flood = yes, Joint = yes, Rockfall = no

        Aval
Bridge    no yes
  high   0.0    
  low    1.0    
  medium 0.0

, , Bearing = bad, Cracks = small, Flood = yes, Joint = yes, Rockfall = no

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = good, Cracks = small, Flood = yes, Joint = yes, Rockfall = no

        Aval
Bridge    no yes
  high   1.0    
  low    0.0    
  medium 0.0

, , Bearing = bad, Cracks = high, Flood = no, Joint = no, Rockfall = yes

        Aval
Bridge    no yes
  high   1.0    
  low    0.0    
  medium 0.0

, , Bearing = good, Cracks = high, Flood = no, Joint = no, Rockfall = yes

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = bad, Cracks = moderate, Flood = no, Joint = no, Rockfall = yes

        Aval
Bridge   no yes
  high      1.0
  low       0.0
  medium    0.0

, , Bearing = good, Cracks = moderate, Flood = no, Joint = no, Rockfall = yes

        Aval
Bridge    no yes
  high   0.0    
  low    1.0    
  medium 0.0

, , Bearing = bad, Cracks = small, Flood = no, Joint = no, Rockfall = yes

        Aval
Bridge   no yes
  high      0.0
  low       0.0
  medium    1.0

, , Bearing = good, Cracks = small, Flood = no, Joint = no, Rockfall = yes

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = bad, Cracks = high, Flood = yes, Joint = no, Rockfall = yes

        Aval
Bridge   no yes
  high      0.0
  low       0.0
  medium    1.0

, , Bearing = good, Cracks = high, Flood = yes, Joint = no, Rockfall = yes

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = bad, Cracks = moderate, Flood = yes, Joint = no, Rockfall = yes

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = good, Cracks = moderate, Flood = yes, Joint = no, Rockfall = yes

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = bad, Cracks = small, Flood = yes, Joint = no, Rockfall = yes

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = good, Cracks = small, Flood = yes, Joint = no, Rockfall = yes

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = bad, Cracks = high, Flood = no, Joint = yes, Rockfall = yes

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = good, Cracks = high, Flood = no, Joint = yes, Rockfall = yes

        Aval
Bridge   no yes
  high      0.0
  low       0.0
  medium    1.0

, , Bearing = bad, Cracks = moderate, Flood = no, Joint = yes, Rockfall = yes

        Aval
Bridge    no yes
  high   0.0    
  low    0.0    
  medium 1.0

, , Bearing = good, Cracks = moderate, Flood = no, Joint = yes, Rockfall = yes

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = bad, Cracks = small, Flood = no, Joint = yes, Rockfall = yes

        Aval
Bridge    no yes
  high   0.0    
  low    1.0    
  medium 0.0

, , Bearing = good, Cracks = small, Flood = no, Joint = yes, Rockfall = yes

        Aval
Bridge   no yes
  high      0.0
  low       1.0
  medium    0.0

, , Bearing = bad, Cracks = high, Flood = yes, Joint = yes, Rockfall = yes

        Aval
Bridge    no yes
  high   0.0    
  low    1.0    
  medium 0.0

, , Bearing = good, Cracks = high, Flood = yes, Joint = yes, Rockfall = yes

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = bad, Cracks = moderate, Flood = yes, Joint = yes, Rockfall = yes

        Aval
Bridge   no yes
  high      0.0
  low       0.0
  medium    1.0

, , Bearing = good, Cracks = moderate, Flood = yes, Joint = yes, Rockfall = yes

        Aval
Bridge   no yes
  high         
  low          
  medium

, , Bearing = bad, Cracks = small, Flood = yes, Joint = yes, Rockfall = yes

        Aval
Bridge    no yes
  high   0.0    
  low    1.0    
  medium 0.0

, , Bearing = good, Cracks = small, Flood = yes, Joint = yes, Rockfall = yes

        Aval
Bridge   no yes
  high         
  low          
  medium

  Parameters of node Cracks (multinomial distribution)

Conditional probability table:

          DTV
Cracks          high       low    medium
  high     0.4000000 0.2500000 0.4285714
  moderate 0.4000000 0.4166667 0.4285714
  small    0.2000000 0.3333333 0.1428571

  Parameters of node DTV (multinomial distribution)

Conditional probability table:
      high       low    medium 
0.3448276 0.4137931 0.2413793

  Parameters of node Earthquake (multinomial distribution)

Conditional probability table:
        no       yes 
0.6206897 0.3793103

  Parameters of node Flood (multinomial distribution)

Conditional probability table:

, , Snow = high

     Rain
Flood      high       low    medium
  no  0.6666667 0.5714286 0.3333333
  yes 0.3333333 0.4285714 0.6666667

, , Snow = low

     Rain
Flood      high       low    medium
  no  0.8000000 0.1666667 0.5000000
  yes 0.2000000 0.8333333 0.5000000

  Parameters of node Joint (multinomial distribution)

Conditional probability table:

     DTV
Joint      high       low    medium
  no  0.5000000 0.2500000 0.5714286
  yes 0.5000000 0.7500000 0.4285714

  Parameters of node Rain (multinomial distribution)

Conditional probability table:
      high       low    medium 
0.2758621 0.4482759 0.2758621

  Parameters of node Rockfall (multinomial distribution)

Conditional probability table:

        Earthquake
Rockfall        no       yes
     no  0.6666667 0.4545455
     yes 0.3333333 0.5454545

  Parameters of node Snow (multinomial distribution)

Conditional probability table:
      high       low 
0.5517241 0.4482759