### 2019-05-05 InfrastructureMaintenance/Programs/IMP-bayesian/IMP-BNsimplifiedexample.tex

Bayesian network parameters
  Parameters of node Avalanche (multinomial distribution)
Conditional probability table:
         Earthquake
Avalanche        no       yes
      no  0.6666667 0.5454545
      yes 0.3333333 0.4545455
  Parameters of node Bridge (multinomial distribution)
Conditional probability table:
, , Rockfall = no
      Avalanche
Bridge        no       yes
   no  0.5000000 0.0000000
   yes 0.5000000 1.0000000
, , Rockfall = yes
      Avalanche
Bridge        no       yes
   no  0.8333333 0.8333333
   yes 0.1666667 0.1666667
  Parameters of node Earthquake (multinomial distribution)
Conditional probability table:
        no       yes 
0.6206897 0.3793103

  Parameters of node Rockfall (multinomial distribution)

Conditional probability table:

        Earthquake
Rockfall        no       yes
     no  0.6666667 0.4545455
     yes 0.3333333 0.5454545

  Bayesian network parameters

  Parameters of node Avalanche (multinomial distribution)

Conditional probability table:

         Earthquake
Avalanche        no       yes
      no  0.6304348 0.5312500
      yes 0.3695652 0.4687500

  Parameters of node Bridge (multinomial distribution)

Conditional probability table:

, , Rockfall = no

      Avalanche
Bridge        no       yes
   no  0.5000000 0.1666667
   yes 0.5000000 0.8333333

, , Rockfall = yes

      Avalanche
Bridge        no       yes
   no  0.7352941 0.7352941
   yes 0.2647059 0.2647059

  Parameters of node Earthquake (multinomial distribution)

Conditional probability table:
        no       yes 
0.5897436 0.4102564

  Parameters of node Rockfall (multinomial distribution)

Conditional probability table:

        Earthquake
Rockfall        no       yes
     no  0.6304348 0.4687500
     yes 0.3695652 0.5312500