### 2019-05-05 InfrastructureMaintenance/General/foreword.tex

%%%%%%%%%%%%%%%%%%%%%%foreword.tex%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\foreword
This document has been compiled to provide students taking the above mentioned course in understanding the examples given in class and assignments. All questions presented have been prepared by the authors based to varying degrees on different sources. The sources are given for each example.

This document is being issued so that students taking the course may better prepare for the exam. As it has not yet undergone the complete rigorous control processes of IBI this document is to be considered a work in progress and is hence labeled as a DRAFT.

\vspace{\baselineskip}
\begin{flushright}\noindent
IBI-ETH, September 2015\hfill {\it Professor: Bryan T. Adey}\\
\hfill {\it Dr: Nam Lethanh}\\
\end{flushright}